---
title: "PERIFOG - power analysis"
author: "Joe M."
format: html
echo: false
---

```{r}
#| label: envir
#| warning: false

library(tidyverse)
library(gt)
library(semTools)
library(presize)
library(semPower)
```

Power analysis for the _"Brain-fog vs. Cognitive Loss in Perimenopause: Evidence-based Assessment"_ **(PERIFOG)** grant proposal.

## Aims:

1. to operationalise Perimenopausal Brain Fog (PMBF) as a multidimentsional latent construct
2. to examine links between PMBF, subjective cognitive complaints (SCC), menopausal symptoms, sleep disturbance, and affective burden
3. to test whether PMBF associates with cognitive profile characterised by reduced cognitive efficiency rather than delayed verbal memory deficit
4. to assess whether PMBF fluctuations are associated with vasomotor symptoms (VMS) burden, autonomic activation and sleep continuity
5. to evaluate whether cognitive and symptom PMBF profile remains after adjustment for post-COVID history

## Variables

```{r}
#| label: tbl-vars

tibble(
  domain = c(
    "(PM) Brain Fog",
    "(PM) Brain Fog",
    "Subjective cognitive complaints",
    "Subjective cognitive complaints",
    "Menopausal symptoms",
    "Menopausal symptoms",
    "Menopausal symptoms",
    "Menopausal symptoms",
    "Sleep",
    "Sleep",
    "Affective burden",
    "Affective burden",
    "Cognition - processing speed",
    "Cognition - processing speed",
    "Cognition - memory",
    "Cognition - memory",
    "Physical activity",
    "Physical activity",
    "Physical activity",
    "COVID"
  ),
  Measure = c(
    "Brain Fog Scale (Atik 2023)",
    "Brain Fog Scale (Debowska 2024)",
    "Everyday Memory Questionnaire-Revised",
    "Le Questionnaire de Plainte Cognitive",
    "Greene Climacteric Scale",
    "COMMA-aligned VMS indicators",
    "cEDA / wearable physiological metrics",
    "Wearable-derived skin temperature",
    "Insomnia Severity Index",
    "Fitbit Sense 2 sleep metrics",
    "State-Trait Anxiety Inventory",
    "Beck Depression Inventory-II",
    "Neuropsychological test",
    "Neuropsychological test",
    "Neuropsychological test",
    "Neuropsychological test",
    "Self-report questionnaire",
    "Senior Fitness Test component",
    "Fitbit Sense 2 accelerometry",
    "Self-report (history + persistent symptoms)"
  )
) |>
  gt()
```

```{r}
#| label: tbl-aims

tibble(
  Aim = c(
    "Aim 1: Operationalise PMBF",
    "Aim 2: Associations (PMBF network)",
    "Aim 3: Cognitive profile of PMBF",
    "Aim 4: Within-person fluctuation links",
    "Aim 5: Adjustment for post-COVID"
  ),
  Outcome = c(
    "PMBF latent construct",
    "PMBF latent construct",
    "Cognitive performance (processing speed, memory)",
    "PMBF (within-person variation)",
    "cognitive outcomes, menopausal symptoms, sleep, affective burden"
  ),
  Exposures = c(
    "BFS+ item pool",
    "SCC, menopausal symptoms, sleep, affective burden",
    "PMBF latent construct",
    "VMS burden, autonomic activation, sleep continuity",
    "PMBF"
  ),
  Covariates = c(
    "None (measurement model)",
    "Age, education, physical activity, COVID history",
    "Age, education, affective burden, sleep, physical activity",
    "Time, individual baseline levels, physical activity",
    "COVID history, age, education, physical activity"
  ),
  Notes = c(
    "CFA + IRT for scale refinement",
    "SEM or network model (possibly bidirectional)",
    "Test dissociation: processing speed vs memory",
    "Multilevel / time-series (EMA or repeated wearable data)",
    "Sensitivity / robustness analysis"
  )
) |>
  gt()
```

## Aim 1

CFA fit via `semTools::findRMSEAsamplesize()` and simulaitons for measurement invariance (BF vs non-BF groups).

## Aim 2

```{r}
#| label: corr-pwr
#| echo: true

# Aim 2
presize::prec_cor(r = seq(.3, .7, .1), conf.width = 0.20)
```

## Aims 3-5

<https://claude.ai/share/313444a2-af9a-47c0-a673-63775fef2b22>

## Aim 3

```{r}
#| label: aim-3
#| echo: true

model.h0 <- "
  PS  =~ 0.707*x1 + 0.707*x2
  MEM =~ 0.707*x3 + 0.707*x4
  PS  ~  0.00*group
  MEM ~  0.00*group
  PS  ~~ 0.00*MEM
"

model.h1 <- "
  PS  =~ 0.707*x1 + 0.707*x2
  MEM =~ 0.707*x3 + 0.707*x4
  PS  ~  0.40*group
  MEM ~  0.00*group
  PS  ~~ 0.00*MEM
"
summary(semPower.powerLav(
    type = "a-priori",
    modelH0 = model.h1,
    modelPop = model.h0,
    alpha = 0.05,
    beta = 0.20
))
```

## Aim 4

```{r}
#| label: aim-4
#| echo: true

# Aim 4
summary(semPower.powerMediation(
    "a-priori",
    bYX = 0.2,
    bMX = 0.2,
    bYM = 0.25,
    alpha = 0.05,
    beta = 0.20,
    nIndicator = c(1, 1, 1),
    loadM = c(1, 1, 1)
))
```


